---
title: "TEMPLATE - alternative chart types practice"
author: "Your Name"
date: 2026-03-04
toc: true
---

::: {.callout-note}
This template accompanies the [alternative chart types practice slides](https://eds-240-data-viz.github.io/course-materials/lecture-slides/SLIDES-alt-charts.html#/title-slide){target="_blank}. Please be sure to cross-reference the slides, which contain important information and additional context!
:::

# Waffle Chart (Sasquatch sightings)

- Typefaces used: Ultra & Josefin Sans
- Colors used: "#357266", "#FFB813", "#983A06", "#005F71", "#757473", "#EFEFEF"

```{r}
#| eval: true
#| echo: true
#| message: false
#| warning: false
#| fig-align: "center"
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                    setup                                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#..........................load packages.........................
library(tidyverse)
library(waffle)
library(showtext)

#..........................import data...........................
bigfoot <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-13/bigfoot.csv')
```


```{r}
#..........................load packages.........................
library(tidyverse)
library(showtext)
library(ggbump)
library(ggtext)

flights <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-07-12/flights.csv') %>% 
  janitor::clean_names()

# Top 6 European countries by total flights
top6 <- flights %>%
  group_by(state_name) %>%
  summarise(total = sum(flt_tot_1, na.rm = TRUE)) %>%
  slice_max(total, n = 6) %>%
  pull(state_name)

# Annual ranks
ranked <- flights %>%
  filter(state_name %in% top6) %>%
  group_by(year, state_name) %>%
  summarise(total = sum(flt_tot_1, na.rm = TRUE), .groups = "drop") %>%
  group_by(year) %>%
  mutate(rank = rank(-total, ties.method = "first")) %>%
  ungroup()

# Blue hues palette
blue_palette <- c(
  "#08306b", "#2171b5", "#4292c6", "#6baed6", "#9ecae1", "#c6dbef"
)

bg_color <- "#3d9e8c"  # greener aquamarine

ggplot(ranked, aes(x = year, y = rank, color = state_name, group = state_name)) +
  geom_bump(linewidth = 1.4, smooth = 8) +
  geom_point(size = 4) +
  geom_text(
    data = filter(ranked, year == min(year)),
    aes(label = state_name), hjust = 1.15, fontface = "bold", size = 3.5
  ) +
  geom_text(
    data = filter(ranked, year == max(year)),
    aes(label = state_name), hjust = -0.15, fontface = "bold", size = 3.5
  ) +
  scale_y_reverse(breaks = 1:6) +
  scale_x_continuous(
    breaks = 2016:2022,
    expand = expansion(mult = c(0.25, 0.25))
  ) +
  scale_color_manual(values = blue_palette) +
  labs(
    title = "European Countries Ranked by Total Flights",
    subtitle = "Top 6 countries by combined incoming & outgoing flights, 2016–2022",
    x = NULL, y = "Rank"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.background = element_rect(fill = bg_color, color = NA),
    panel.background = element_rect(fill = bg_color, color = NA),
    panel.grid.major = element_line(color = "white", linewidth = 0.3, linetype = "dashed"),
    panel.grid.minor = element_blank(),
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 11, margin = margin(b = 10)),
    legend.position = "none",
    plot.margin = margin(20, 20, 20, 20)
  )

```

# Bump chart (highest paying occupations)

- Typefaces used: Passion One & Oxygen
- Colors used: `viridisLite::magma()` & "#2A114E", "#6D6B71", "#FFF8F4"

```{r}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                    setup                                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#..........................load packages.........................
library(tidyverse)
library(ggbump)
library(ggtext)
library(showtext)

#..........................import data...........................
jobs <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/jobs_gender.csv")
```

# Stream graph (energy production)

- Typefaces used: Exo 2 & Open Sans
- Colors used: `paletteer::paletteer_d("MetBrewer::Archambault", n = 7)`

```{r}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                    setup                                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#..........................load packages.........................
library(tidyverse)
library(ggstream)
library(ggtext)
library(showtext)

#..........................import data...........................
capacity <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2022/2022-05-03/capacity.csv")
```











